### 2017-03-31 paper/evaluation.tex
pass on text

According to Maven's documentation \Fix{cite surefire doc}, any
parallelization setting depends either on (1) \CodeIn{parallel} to
define the parallelism level within a JVM followed by
\CodeIn{threadCounter} or (2) \CodeIn{forkCount} (or \CodeIn{forkMode}
in older versions) to define the number of forked JVMs. For each test
execution, we look for theses parameters and their values to classify
them.  \Jbc{Missing: explain how we mapped the configurations to
parallel level.} Figure~\ref{fig:freqmodes-dynamic} summarizes our
findings.

>>>>

According to Maven's documentation~\cite{maven-surefire-plugin}, any
parallelization setting depends either on (1) \CodeIn{parallel} to
define the parallelism level within a JVM followed by
\CodeIn{threadCounter} or (2) \CodeIn{forkCount} (or \CodeIn{forkMode}
in older versions) to define the number of forked JVMs. For each test
execution, we look for theses parameters and their values to classify
them.  \Jbc{Missing: explain how we mapped the configurations to
parallel level.} Figure~\ref{fig:freqmodes-dynamic} summarizes our
findings.

### 2017-03-31 paper/modes.tex
pass on text

\newcommand{\Seq}{L0}
\newcommand{\SeqClassParMeth}{L1}
\newcommand{\ParClassSeqMeth}{L2}
\newcommand{\ParClassParMeth}{L3}
\newcommand{\Fork}{F}
\newcommand{\ForkSeq}{\Fork{}\Seq{}}
\newcommand{\ForkParMeth}{\Fork{}\SeqClassParMeth{}}

@@

%\Mar{add a sentence to introduce this section at a higher abstraction
%  level.}

This section provides some background on parallel test execution.

Parallelism in test execution can be obtained at different levels.
Figure~\ref{fig:levels} illustrates relevant levels.  The highest
level indicates parallelism that can be obtained through different
machines on the network.  For instance, using virtual machines from a
cloud service to perform a distributed execution.  The machine space
sits under the ``network space'' and provides a \emph{lower-level}
parallelism.  In this case, computation can be offloaded at different
CPUs within a machine and at different threads within each CPU.  Note
that all these levels are complementary:  low-level parallelization
schemes could leverage the computing power of server nodes in addition
to the aggregate processing power of the farm.

%% Lower-level parallelism fits
%% particularly well smaller organizations (/projects) with relatively
%% high testing costs but lower budgets.

@@

\Comment{
    Unfortunately, these solutions can't be used out-of-the-box:
    unrestricted parallel execution of tests can produce
    non-deterministic results as developers typically do not provision
    protection to concurrent accesses originated from arbitrary program
    points (\ie{}, tests).  We refer to this problem as Parallel
    Execution Flakiness (\pef{}).

    %% These solutions enable the
    %% use of commodity hardware to maximize CPU usage\footnote{In the case
    %%   of the Java language, for example, it is possible to explore
    %%   parallelism across and within JVMs.}.

    To illustrate the importance of parallel test execution and
    the problem of \pef{} let's consider the case of the ``core'' module
    from the Apache Camel project~\cite{apache-camel-web}.  This module
    contains 5,679 test cases, declared in 2,356 test classes.  We ran
    those tests in a machine with 16GB of memory and 8 virtual CPUs (4
    cores with 2 native threads each).  Sequential test execution takes
    24m50s to run this test set.  Execution of the same test set takes
    2m28s when we configured parallel execution to fork a JVM per CPU and
    execute test classes, uniquely allocated to that JVM, sequentially but
    running test methods from each class in separate threads.  Note that
    this is an order of magnitude speedup (10.07x)\Fix{Need to understand
    why this is 10x as opposed to something closer to 7x - I didn't get
    your concern here}.  Unfortunately, due to \pef{}, $\sim$2\% (114 of
    5,679) of the tests fail when executed in parallel.  It is important
    to notice that the ratio of failures varies with the project as it
    depends on factors such as length of test cases and amount of shared
    state across tests.

    \pef{} is an important obstacle to enable parallel test execution.
    Conceptually, higher parallelization can result in higher chances of
    concurrency-related problems.  It is important to execute tests
    efficiently without sacrificing reliability
    \Fix{gap between pars?? What gap?}
}

%%Build systems delegate functionalities to the testing framework they
%%use~\cite{maven-surefire-plugin}.  Figure~\ref{fig:surefire}
%%illustrates how to configure Maven for mode
%%\ParClassSeqMeth{}~--~classes sequentially (default) and test methods
%%in parallel. Although this configuration is defined in the build file
%%(\eg, \emph{pom.xml} on Maven), these settings are forwarded to the
%%underlying testing framework (in this case, JUnit).

%% \Jbc{I have to understand better how Surefire executes tests. I'm
%% afraid it has its own wrapper classes and reuse JUnit core
%% functionalities. I'm saying this because JUnit does not offer a finer
%% control over thread numbers while this configuration is possible with
%% Surefire} For example,

%\Fix{we
%  need to know how Maven allocates test classes to JVMs}

@@

The list below shortly describes the choices to control parallelism
within a Java virtual machine (JVM).

\begin{itemize}
\item
    \textbf{Sequential (\Seq).}~This configuration corresponds to the
        default behaviour in a test execution. No parallelism is
        involved.
\item
    \textbf{Sequential classes; parallel methods
        (\SeqClassParMeth).}~When running a test class, the testing
        framework executes methods on multiple threads. Each class
        runs sequentially.
\item
    \textbf{Parallel classes; sequential methods
        (\ParClassSeqMeth{}).}~The testing framework runs test classes
        on multiple threads and methods run sequentially within each
        thread.
\item
    \textbf{Parallel classes; Parallel methods
        (\ParClassParMeth).}~This configuration corresponds
        conceptually to the union of \ParClassSeqMeth{} and
        \SeqClassParMeth{}: the testing framework runs test methods
        from several classes on multiple threads.
\end{itemize}

Typically, testing frameworks executes tests in sequence by default
(\Seq{}) and may allow the developer to use an ordering criteria (\eg,
test name).  For the \SeqClassParMeth{} setting, \Fix{elaborate...}.
For the \ParClassSeqMeth{} setting, \Fix{elaborate...}. Finally, the
\ParClassParMeth{} setting \Fix{...elaborate...}.

@@

The list below shortly describes the choices to control parallelism
through build systems:

\begin{itemize}
\item
    \textbf{Forked classes; sequential methods (\ForkSeq).}~This
        configuration corresponds to the parallel execution of test
        classes on different JVMs and test methods run sequentially
        within each JVM.
\item
    \textbf{Forked classes; parallel methods (\ForkParMeth).}~This
        configuration corresponds to the parallel execution of test
        classes on different JVMs and test methods run on multiple
        threads within each JVM.
\end{itemize}

Several build systems support the execution of test classes in forked
processes running a JVM. In the \ForkSeq{} configuration, each spawned
process runs a different test class at time and the builder merges the
results from each execution. In the \ForkParMeth{} configuration, the
build system forwards settings to the underlying testing framework to
enable parallel execution within each process in addition to running
test classes on different processes.  However, some configurations
(\eg, \ParClassSeqMeth) may not take effect since only one test class
can run in a forked process at time.  To the best of our knowledge, we
are unaware of any build system capable of running multiple classes at
time within a forked process.  \Comment{In addition, the builder can
reuse these processes to execute the next test classes in the queue
according to the developer's preference. For instance, one might want
the builder to spawn a new process for each test class to ensure
maximum isolation. Notice that to exploit parallelism at build system
level, the number of forked processes must be proportional to the
number of available CPUs, otherwise, the high frequency of
context-switching and processes competing for a CPU may degrade the
execution performance  \Jbc{$\leftarrow$ I'm not sure if we should
keep this. We can remove it if necessary}.}

%% In this paper, we use the prefix ``\Fork{}'' to indicate that a given
%% testing configuration has the feature of spawning JVMs to run test
%% classes enabled.

\subsubsection*{Illustration}~\Fix{elaborate the following example to
conclude this section}

\begin{figure}[h!]
\centering
\scriptsize
\lstset{
    escapeinside={@}{@},
    numbers=left,xleftmargin=1em,frame=single,framexleftmargin=0.5em,
    basicstyle=\ttfamily\scriptsize, boxpos=c, numberstyle=\tiny,
    morekeywords={parallel, threadCount, perCoreThreadCount,
    forkCount, reuseFork},
    deletekeywords={true}
}
\begin{lstlisting}
<plugin>
    <groupId>org.apache.maven.plugins</groupId>
    <artifactId>maven-surefire-plugin</artifactId>
    <configuration>
        <forkCount>1C</forkCount>
        <parallel>methods</parallel>
        <threadCount>5</threadCount>
    </configuration>
</plugin>
\end{lstlisting}
    \caption{\label{fig:surefire} Maven with a \ForkParMeth{}
    configuration.  Maven forks one JVM per core (\CodeIn{forkCount}
    parameter) and uses five threads (\CodeIn{threadCount} parameter)
    to run methods (\CodeIn{parallel} parameter) within each JVM.}

%%    test methods run in parallel (\CodeIn{parallel}
%%    parameter) using two threads (\CodeIn{threadCount} parameter) per
%%    available core (flag \CodeIn{perCoreThreadCount}).}
\end{figure}

%% \Fix{-----}
%% In JUnit, \emph{ParallelComputer} provides support to parallel
%% execution: it instantiates a test runner with an
%% \emph{ExecutorService} from the Java Concurrent API. Each test method
%% is executed in a separated thread as the thread pool creates new
%% threads on-demand and reuse them if any thread is available.
%% \Comment{http://junit.org/junit4/javadoc/latest/src-html/org/junit/experimental/ParallelComputer.html#line.14}

%% As in \ParClassSeqMeth, the test runner has a cached thread poll but
%% instead of executing test methods from a single test class, the test
%% suite runner executes classes in separated threads and methods
%% sequentially within each thread.

%% \Jbc{I need to confirm this but AFAIK this is how it happens} As in
%% \SeqClassParMeth, classes execute on separated threads but test
%% methods are also executed in parallel. More precisely, there is a
%% thread pool to execute classes in separated threads (\ie, like in
%% \ParClassSeqMeth) and, within each class runner, there is an
%% \emph{ExecutorService} to run test methods in parallel (\ie, like in
%% \SeqClassParMeth).

%% \Fix{-----}

>>>>

\newcommand{\Seq}{C0}
\newcommand{\SeqClassParMeth}{C1}
\newcommand{\ParClassSeqMeth}{C2}
\newcommand{\ParClassParMeth}{C3}
\newcommand{\Fork}{F}
\newcommand{\ForkSeq}{\Fork{}\Seq{}}
\newcommand{\ForkParMeth}{\Fork{}\SeqClassParMeth{}}

@@

This section provides important background on parallel test execution.

Parallelism in test execution can be obtained at different levels.
Figure~\ref{fig:levels} illustrates these levels.  The highest
level indicates parallelism that can be obtained through different
machines on the network.  For instance, using virtual machines from a
cloud service to perform a distributed execution.  The machine space
sits under the ``network space'' and provides a \emph{lower-level}
parallelism.  In this case, computation can be offloaded at different
CPUs within a machine and at different threads within each CPU.  Note
that all these levels are complementary:  low-level parallelization
schemes could leverage the computing power of server nodes in addition
to the aggregate processing power of the farm.

@@

@@

The list below shows the choices to control parallelism within a Java
virtual machine (JVM) which runs in a single process of the CPU.

\begin{itemize}
\item
    \textbf{Sequential (\Seq).}~No parallelism is involved.
\item
    \textbf{Sequential classes; parallel methods
        (\SeqClassParMeth).}~This configuration corresponds to the
        sequential execution of test classes and test methods run in
        multiple threads.
\item
    \textbf{Parallel classes; sequential methods
        (\ParClassSeqMeth{}).}~This configuration corresponds to the
        execution of test classes in multiple threads and tests run
        sequentially within each thread.
\item
    \textbf{Parallel classes; Parallel methods
        (\ParClassParMeth).}~This configuration corresponds
        conceptually to the union of \ParClassSeqMeth{} and
        \SeqClassParMeth{}: the testing framework runs test methods
        from several classes on multiple threads.
\end{itemize}

The configuration \Seq{} is a proper fit for short-running test suites
since this is the default configuration and the execution provides
feedback in an acceptable time. However, for costly suites, it is
impractical for developers to wait long-running executions to take
action.  In this scenario, the remaining configuration may help to
amortize the execution cost depending on how the test suite was
designed. For instance, the configuration \SeqClassParMeth{} may
speedup the execution of test suites composed large test classes since
the cost of running all test methods is amortized by a pool of
threads. Similarly, the configuration \ParClassSeqMeth{} may speedup
test suites composed by several small test classes. Notice that both
configurations \ParClassSeqMeth{} and \SeqClassParMeth{} avoid
data-race condition at different levels. For instance, if two test
classes performs I/O operations to the same file, the configuration
\SeqClassParMeth{} would be preferable. Finally, the configuration
\ParClassParMeth{} does not impose any sequential execution but it is
more likely to manifest data-races as methods from different classes
run in parallel.

@@

The list below shows the choices to control parallelism through build
systems which uses multiple JVMs on different processes each:

\begin{itemize}
\item
    \textbf{Forked classes; sequential methods (\ForkSeq).}~This
        configuration corresponds to the parallel execution of test
        classes on different JVMs with test methods running
        sequentially within each JVM.
\item
    \textbf{Forked classes; parallel methods (\ForkParMeth).}~This
        configuration corresponds to the parallel execution of test
        classes on different JVMs with test methods running on
        multiple threads within each JVM.
\end{itemize}

Several build systems support the execution of test classes in forked
processes running a JVM. In the configuration \ForkSeq{}, each spawned
process runs a different test class at time and the builder merges the
results from each execution. In the configuration \ForkParMeth{}, the
build system forwards settings to the underlying testing framework to
enable parallel execution within each process in addition to running
test classes on different processes.  However, some configurations
(\eg, \ParClassSeqMeth) may not take effect since only one test class
can run in a forked process at time.  To the best of our knowledge, we
are unaware of any build system capable of running multiple classes at
time within a forked process.

\subsubsection*{Illustration}~Figure~\ref{fig:surefire} illustrates
the support of parallel execution with the configuration
\ForkParMeth{} on Maven, a popular build system in Java projects.
Maven uses a configuration file based on XML (known as \pomf{}) to
define tasks related to the project building and these tasks are
performed several plugins. \Jbc{falar do surefire e explicar a
configuracao ilustrada - Surefire levanta uma JVM por core e cria uma
pool de 5 threads em cada JVM para executar testes em paralelo.}

\begin{figure}[h!]
\centering
\scriptsize
\lstset{
    escapeinside={@}{@},
    numbers=left,xleftmargin=1em,frame=single,framexleftmargin=0.5em,
    basicstyle=\ttfamily\scriptsize, boxpos=c, numberstyle=\tiny,
    morekeywords={parallel, threadCount, perCoreThreadCount,
    forkCount, reuseFork},
    deletekeywords={true}
}
\begin{lstlisting}
<plugin>
    <groupId>org.apache.maven.plugins</groupId>
    <artifactId>maven-surefire-plugin</artifactId>
    <configuration>
        <forkCount>1C</forkCount>
        <parallel>methods</parallel>
        <threadCount>5</threadCount>
    </configuration>
</plugin>
\end{lstlisting}
    \caption{\label{fig:surefire} Maven setup with the configuration
    \ForkParMeth{}.  Maven forks one JVM per core (\CodeIn{forkCount}
    parameter) and uses five threads (\CodeIn{threadCount} parameter)
    to run methods (\CodeIn{parallel} parameter) within each JVM.}
\end{figure}
